---
title: "DATA607Assignment4"
author: "Erick Hadi"
date: "2024-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup and load libraries and data

Install and load in the neccesary libraries 
```{r get-libraries}
library(tidyverse)
library(openintro)
library(dplyr)
library(ggplot2)
```

Load the data into a data frame and display
```{r load-data}
flights_data_wide <- read.csv('https://raw.githubusercontent.com/ErickH1/DATA607Assignment4/refs/heads/main/data607hw4data.csv')
```
```{r load-data-head}
head(flights_data_wide)
glimpse(flights_data_wide)
```
# Tidy Data
Tidying up the data by removing na values 
```{r remove-na-values-rename-column-names}
flights_data_wide <- flights_data_wide %>%
  drop_na() %>% 
  rename(airline = X, status = X.1)
```

```{r fill-na}
flights_data_wide <- flights_data_wide %>%
  mutate(airline = na_if(airline, '')) %>% 
  fill(airline, .direction = c('down'))
```

Making the data frame long
```{r data-frame-long}
flights_data_long <- flights_data_wide %>%
  pivot_longer(cols = c("Los.Angeles", "Phoenix", "San.Diego", "San.Francisco", "Seattle"),
               names_to = "city",
               values_to = "count")
head(flights_data_long)
```
Finally tidying the data to the desired dataframe
```{r data-frame-tidy}
flights_data_tidy <- flights_data_long %>%
  pivot_wider(names_from = status,
              values_from = count)

names(flights_data_tidy)[3] <- "on_time"
head(flights_data_tidy)
```

# Data Analysis
Adding delay and on time rate as their own columns
```{r data-tidy-delayed-on_time-rate}
flights_data_tidy <- flights_data_tidy %>%
  mutate(delayed_rate = delayed / (on_time + delayed))

flights_data_tidy <- flights_data_tidy %>%
  mutate(ontime_rate = on_time / (on_time + delayed))

glimpse(flights_data_tidy)
```
Create and display plots
```{r plot-delayed-rate}
ggplot(flights_data_tidy, aes(fill= airline, y = delayed_rate, x = city)) +
  geom_bar(position="dodge", stat = "identity")
```
```{r plot-ontime-rate}
ggplot(flights_data_tidy, aes(fill= airline, y = ontime_rate, x = city)) +
  geom_bar(position="dodge", stat = "identity")
```

# Conclusion
In conclusion, based on the analysis on the two airlines, AM West has a higher delayed rate in every arrival in a city. 








